Inference of proto-neutron star properties from gravitational wave data in core-collapse supernovae.

### 2020-09-29 introduction.tex
Update

{In this work, we present a method to infer from the GW data alone, the time evolution of some properties or the PNS, namely a combination of its mass and radius. For this purpose we have developed an algorithm to  
extract the time-frequency evolution of the main feature in the spectrograms of the GW emission of 2D simulations of CCSN. This feature corresponds to the $^2\rm{g}_2$ mode, according to the nomenclature used in \cite{Torres:2019b} (different authors may have slightly different naming convection). Next, we use the universal relations obtained by \cite{Torres:2019b}{, based on a set of 1D simulations,} to infer the time evolution of the ratio $M_{\rm PNS}/R_{\rm PNS}^2$, being 
$M_{\rm PNS}$ and $R_{\rm PNS}$ the mass and radius of the PNS.} Using 2D CCSN waveform corresponding to different progenitor masses we estimate the performance of the algorithm for current and future generation of ground-based GW detectors.

This paper is organised as follows. Section II describes the details of the CCSN simulations used in the paper. Section III focuses on the algorithm that extracts the time evolution of a combination of the mass and radius of the PNS corresponding to a g-mode. Section IV shows the performance of the data analysis method for different GW detectors. Finally, we discuss the results in section V.

>>>>

{In this work, we present a method to infer from the GW data alone, the time evolution of some properties or the PNS, namely a combination of its mass and radius. For this purpose we have developed an algorithm to  
extract the time-frequency evolution of the main feature in the spectrograms of the GW emission of 2D simulations of CCSN. This feature corresponds to the $^2\rm{g}_2$ mode, according to the nomenclature used in \cite{Torres:2019b} (different authors may have slightly different naming convention). Next, we use the universal relations obtained by \cite{Torres:2019b}{, based on a set of 1D simulations,} to infer the time evolution of the ratio $M_{\rm PNS}/R_{\rm PNS}^2$, being 
$M_{\rm PNS}$ and $R_{\rm PNS}$ the mass and radius of the PNS.} Using 2D CCSN waveform corresponding to different progenitor masses we estimate the performance of the algorithm for current and future generation of ground-based GW detectors.

This paper is organised as follows. Section II describes the details of the CCSN simulations used in the paper. Section III focuses on the algorithm that extracts the time evolution of a combination of the mass and radius of the PNS corresponding to a g-mode. Section IV shows the performance of the data analysis method with simulated GW detectors data. Finally, we discuss the results in section V.

### 2020-09-29 results.tex
Update

For each of the noise realisations, we reconstruct the ratio time series {$r_i$}
of length $N$ starting from the left side of the spectrogram and constraining the beginning of the
track to be smaller than \unit[200]{Hz}. The reconstructed ratio is then compare to the ``true'' ratio
{$r_i^0$} derived from the PNS mass and radius \pcd{computed from the \sout{generated by the simulation code that produced} {\texttt s20S} simulation}.

Figure \ref{fig:s20results} is showing the fraction of the ratio {$r_i^0$} values that fall
within the 95\% confidence interval of {$r_i$}. This quantity, $coverage$, is taking maximal values
when the source is located within few kpc and then decreases with the distance.

@@

$\Delta$ values of each of the 100 noise realisation are shown as well as function of the distance
on Figure \ref{fig:s20results}. For a source located up to $\sim$\unit[10]{kpc} the relative error
remains smaller than 20\% \pcd{[PCD: should we comment on the outliers?]}. At small distance $\Delta$ is small but not null. This reflects the
approximation of the model used for $r$.
It is nevertheless remarkable that one can reconstruct the ratio time series with a good
precision at distance up to $\sim$ 10 kpc for this particular waveform, with $coverage$ value
larger than 80\%.
We have tested that the method does not depend on features of {\texttt s20S} using \pcd{the} 7 other waveforms \pcd{of the {\it test set}}
described in section \ref{sec:simulations} covering a large range of progenitor masses.
Figure \ref{fig:aLIGO_cov_allwvf} shows that apart {\tt s11.2--LS220}, the ratio is well
reconstructed for all waveforms up to $\sim$ 10kpc. On this figure we also show the $coverage$
value in case of absence of signal. The median value is significantly different from 0 because
the g-mode reconstruction algorithm is looking for a continuously frequency increasing track
in the spectrogram, starting between 0 and 200 Hz. In Table \ref{tab:results} we are reporting the
distance at which $coverage$ median is lower than 80\%.

@@

@@

\hline
\multirow{2}{*}{aLIGO} & $d_{max}$ (kpc) & 7.3  & 26.3 & 24.8 & 22.3 & 16.3 & 10.8 & 7.8  & 14.8\\
\cline{2-10}
                       & SNR            & 13.0 & 37.0 & 38.7 & 39.2 & 22.9 & 23.3 & 75.6 & 63.6\\
\hline
\hline

  \end{tabular}
  \caption{%%
    Matched filter signal-to-noise ratio (SNR) of the simulated waveforms for
    the different GW detectors considered in this study. The source is located
    at 10 kpc and is optimally oriented with respect to the detector.
    %%
  }
  \label{tab:results}
\end{table}

@@

>>>>

For each of the noise realisations, we reconstruct the ratio time series {$r_i$}
of length $N$ starting from the left side of the spectrogram and constraining the beginning of the
track to be smaller than \unit[200]{Hz}. The reconstructed ratio is then compare to the ``true'' ratio
{$r_i^0$} derived from the PNS mass and radius computed from the {\texttt s20S} simulation.

Figure \ref{fig:s20results} is showing the distribution of the fraction of the ratio {$r_i^0$} values that fall
within the 95\% confidence interval of {$r_i$}. This quantity, $coverage$, is taking maximal values
when the source is located within few kpc and then decreases with the distance.

@@

$\Delta$ values of each of the 100 noise realisation are shown as well as function of the distance
on Figure \ref{fig:s20results}. For a source located up to $\sim$\unit[10]{kpc} the relative error
remains smaller than 20\% \pcd{[PCD: should we comment on the outliers?]}. At small distance
$\Delta$ is small but not null. This reflects the approximation of the model used for $r$.
It is nevertheless remarkable that one can reconstruct the ratio time series with a good
precision at distance up to $\sim$ 9 kpc for this particular waveform, with $coverage$ value
larger than 80\%.
We have tested that the method does not depend on features of {\texttt s20S} using \pcd{the} 7
other waveforms \pcd{of the {\it test set}} described in section \ref{sec:simulations} covering
a large range of progenitor masses.
Figure \ref{fig:aLIGO_cov_allwvf} shows that apart {\tt s11.2--LS220}, the ratio is well
reconstructed for all waveforms up to $\sim$ 9kpc.
To measure the level of contamination of $coverage$ by reconstructed ratios not associated to a
signal, we have run 100 simulations without injecting a signal. The median of $coverage$ as well as
the 95 quartile are shown on Figure \ref{fig:aLIGO_cov_allwvf}. The noise only median value is
significantly different from zero because the g-mode reconstruction algorithm is looking for a
continuously frequency increasing track in the spectrogram, starting between 0 and 200 Hz.
Outliers can also reach values as high as 80\%.
Figure \ref{fig:aLIGO_prec_allwvf} shows $\Delta$ as function of the distance for the same signals
as well as the result when only noise is considered.
In Table \ref{tab:results} we are reporting the distance at which $coverage$ median is lower than
95\% of the noise only values. These numbers are an estimate of the order of magnitude of the 
source maximal distance at which a reconstruction of the ratio could be possible with current
and future generations of gravitational wave detectors. We have checked that $coverage$ and $\Delta$
provide similar values.

\mab{add comment about ADV}

@@

The same analysis has been performed using expected sensitivity curves for the third generation of
gravitational wave detectors. The US based project Cosmic Explorer~\cite{} is proposing to reach out
its design sensitivity circa 2040 through two phases labeled CE1 and CE2 also shown in
Figure~\ref{fig:spectrum}. In Europe, the Einstein Telescope is composed of three nested interferometers
in a configuration of an equilateral triange~\cite{}.

\mab{explain ET config}
\mab{comment figure 9}

@@

\hline
\multirow{3}{*}{aLIGO} & $d_{r}$ & 7   & 27  & 24  & 22  & 16  & 10  & 37  & 45 \\
\cline{2-10}
                       & $d_{det}$ & 11 & 36 & 26 & 27 & 21 & 16 & 74 & 61\\
                       & SNR      & 14 & 46 & 33 & 35 & 27 & 21 & 96 & 80\\

\hline
\hline
\multirow{3}{*}{ADV}   & $d_{r}$ & 6   & 25  & 20  & 19  & 14  & 10  & 43  & 42 \\
\cline{2-10}
                       & $d_{det}$ &  &  &  &  &  &  &  & \\
                       & SNR      &  &  &  &  &  &  &  & \\

\hline
\hline
\multirow{2}{*}{CE1}   & $d_{r}$   & 79  & 303 & 257 & 229 & 186 & 115 & 523 & 490 \\
\cline{2-10}
                       & $d_{det}$ & 115 & 377 & 270 & 282 & 217 & 168 & 774  & 633\\
                       & SNR      & 149 & 490 & 352 & 366 & 282 & 218 & 1006 & 822\\

\hline
\multirow{2}{*}{CE2}   & $d_{r}$  & 135 & 518 & 464 & 405 & 305 & 183 & 391 & 700 \\
\cline{2-10}
                       & $d_{det}$ & 197 & 649 & 468 & 489 & 375 & 294 & 1347  & 1100\\
                       & SNR    & 256 & 843 & 608 & 635 & 487 & 382 & 1751 & 1430\\

\hline
\multirow{2}{*}{ET\_B} & $d_{r}$ & 71  & 293 & 248 & 245 & 158 & 123 & 113 & 392 \\
\cline{2-10}
                       & $d_{det}$ & 106 & 364 & 274 & 391 & 216 & 200 & 805 & 665\\
                       & SNR      & 138 & 473 & 356 & 379 & 381 & 260 & 1046 & 865\\

\hline
\multirow{2}{*}{ET\_C} & $d_{r}$ & 75  & 302 & 239 & 237 & 172 & 131 & 239 & 446 \\
\cline{2-10}
                       & $d_{det}$ & 97 & 332 & 246 & 260 & 194 & 164 & 727  & 603\\
                       & SNR      & 126 & 432 & 320 & 338 & 252 & 213 & 945 & 783\\

\hline
\multirow{2}{*}{ET\_D} & $d_{r}$ & 83  & 329 & 257 & 261 & 186 & 139 & 369 & 523 \\
\cline{2-10}
                       & $d_{det}$ & 107 & 368 & 271 & 285 & 213 & 174 & 796  & 661\\
                       & SNR      & 140 & 477 & 352 & 371 & 277 & 227 & 1034 & 859 \\

  \end{tabular}
  \caption{%%
    Maximal distance $d_{max}$ at which the ratio $r=M_{\rm PNS}/R_{\rm PNS}^2$ is reconstructed
    with good accuracy for a source optimally oreinted with respect to the GW detectors
    considered in this study.
    Matched filter signal-to-noise ratio (SNR) of the simulated waveforms for
    the different GW detectors. The source is located at 10 kpc and is optimally oriented
    with respect to the detector.
    %%
  }
  \label{tab:results}
\end{table}

@@

\begin{figure}
  \centering
  \includegraphics[width=0.5\textwidth]{plots/dist_allwvfs_2G3G}
  \caption{} \label{fig:distances}
\end{figure}
